### 2023-04-27 paper/main.tex

\documentclass{article}

\usepackage{url}
\usepackage{todonotes}
\usepackage{hyperref}

\begin{document}

\title{An Analysis of the Security Blind Spots of Software Composition Analysis caused By Cloning and Shading }

\maketitle

\abstract{
Modern software heavily relies on the use of components. Those components are usually published  in central repositories, and managed by build systems via dependencies. Due to issues around vulnerabilities, licenses and the propagation of bugs, the study of those dependencies is of upmost importance, and various software composition analysis tools have emerged to deal with those challenges.

A particular challenge those tools phase is obfuscated dependencies that are the result of cloning or shading where code from a component is "inlined" into a client, and sometimes moved into a different namespace.

We study the practice of shading, and how prevalent this is in the Maven ecosystem used by Java and other JVM languages (Kotlin, Scala, etc). We also present a tool pipeline to detected cloned and shaded artifacts in the Maven repository.  Our tool is lightweight in that it does not requires the creation and maintenance of an index, and uses a custom AST-based clone detection.

We evaluates this by detecting the presence of known vulnerabilities with assigned CVEs in artifacts not flagged by mainstream SCA tools. Starting with TODO known CVEs and proof-of-concept projects demonstrating the presence of the respective vulnerabilities  in the respective artifacts, we query the Maven repository for XX potential clones, and detect TODO other artifacts with those vulnerabilities. Our analysis is precise by design.

}

\input{introduction}
\input{relatedwork}
\input{detection}
\input{evaluation}
\input{conclusion}

\section{Acknowledgements}

\todo[inline] {Dhanushka, Emanual, Oracle, SFTI}

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}